Dynamic Walking and Whole-Body Motion Planning for Humanoid Robots: an Integrated Approach

### 2012-12-04 paper.tex
Use balance term insteand of stability for paths and trajectories.

This paper presents a general method for planning collision-free whole-body walking motions
for humanoid robots. First, we present a randomized algorithm for constrained motion
planning, that is used to generate collision-free statically stable paths solving
manipulation tasks. Then, we show that dynamic walking makes humanoid robots 
small-space controllable. Such a property allows to easily transform 
collision-free statically balanced paths into 
collision-free dynamically balanced trajectories. It leads to a sound 
algorithm which has been applied and evaluated on several
problems where whole-body planning and  walk are needed,
and the results have been validated  on a real HRP-2 robot.

@@

In this paper, we present a planning algorithm that considers exact models of a humanoid
robot and its environment. It is used to solve navigation and manipulation problems. Our planner 
generates collision-free, statically stable 
geometric paths. We also present a constructive formal proof that these paths can be approximated by
collision-free, dynamically balanced walk trajectories. Our planner is designed for indoor
environments, where the floor is horizontal and flat. Because we do not compute explicitly footstep
positions at the planning stage, the presented planner is unable to plan motions in which the
robot steps over obstacles. This limitation is discussed in the paper.

@@

Another  field of  humanoid  robotics research  is  the generation  of
dynamically balanced walk patterns. Since  the introduction of  the ZMP
formalism  \cite{vukobratovic1969contribution},  several  methods  have  been  proposed  to
generate  walking  motions efficiently.   One  way  to  deal with  the
complexity  of a  humanoid  robot  kinematic tree  is  to use  the
so-called "cart-table" simplified model \cite{kajita2003biped}. Based on this
model,  planning a trajectory  for the  ZMP is  reduced to  planning a
trajectory  for  the Center  of  Mass (CoM)  of  the  robot.  Given  a
trajectory  of  the CoM  and  footprint  positions, inverse  kinematics
solvers can animate  the whole set of DoFs of the
robot to generate a dynamically stable walk trajectory.

@@

\section{From Statically Stable Paths to Dynamic Walk Trajectories}
\label{sec:wb-step}

@@

\paragraph{Use of ZMP preview controller:} The control strategy presented in the  
previous proof may
generate very  long trajectories, because  of the transient  states at
the beginning and end of the locomotion. In the actual implementation,
we have chosen  to generate  CoM motions with  a ZMP preview  controller, as
presented in \cite{kajita2003biped}.  We  have observed experimentally that the
amplitude of  CoM trajectories decreases  when the frequency  of steps
increases. Our current ZMP preview controller relies on the cart-table model approximation.
To make this approximation valid, we fix the height of the robot CoM during walk,
as well as the vertical orientation of the robot waist. These geometric constraints 
are also applied when planning statically stable paths, to ensure that the paths
can be approximated by dynamic walk trajectories. Note that this is
due to our current  ZMP preview controller implementation, and does not affect
the generality of the small-space controllability result presented above.

@@

\subsection{Application: Dynamic Approximation of a Statically Stable Sliding Path}
\label{sec:ssc-application}

The  algorithm   that  animates  a  statically  stable   path  into  a
dynamically stable  walk trajectory has been inspired  by the previous
small-space controllability  proof.
Given a statically  stable path $p$ respecting the cart-table model approximation constraints,
we  start  by placing  footprints  corresponding  to  the nominal  walk
pattern  of the  robot. Given the footprints, we  compute a ZMP trajectory, 
and a preview controller outputs  a corresponding CoM trajectory. 
The stack of tasks applied to
the robot to generate a dynamic walk motion is -- in decreasing priority order:

@@

Tasks (1)  and (2) generate a  dynamically stable motion  by using the
simplified cart-table model  and the ZMP criterion. Tasks  (3) and (4)
ensure that the  resulting motion is well described  by the cart-table
model. Task (5)  is used to approximate $p$ as  well as possible given
the walk parameters.

>>>>

This paper presents a general method for planning collision-free whole-body walking motions
for humanoid robots. First, we present a randomized algorithm for constrained motion
planning, that is used to generate collision-free statically balanced paths solving
manipulation tasks. Then, we show that dynamic walking makes humanoid robots 
small-space controllable. Such a property allows to easily transform 
collision-free statically balanced paths into 
collision-free dynamically balanced trajectories. It leads to a sound 
algorithm which has been applied and evaluated on several
problems where whole-body planning and  walk are needed,
and the results have been validated  on a real HRP-2 robot.

@@

\textbf{In this paper, we present a planning algorithm that considers exact models of a humanoid
robot and its environment. It is used to solve navigation and manipulation problems. Our planner 
generates collision-free, statically balanced
geometric paths, i.e. paths such that for any path configurations, the robot is balanced on its
feet with zero speed. We also present a constructive formal proof that these paths can be
approximated by collision-free, dynamically balanced walk trajectories, i.e. trajectories such
that for any trajectory instant, the robot is balanced on at least one foot.} Our planner is designed for indoor
environments, where the floor is horizontal and flat. Because we do not compute explicitly footstep
positions at the planning stage, the presented planner is unable to plan motions in which the
robot steps over obstacles. This limitation is discussed in the paper.

@@

Another  field of  humanoid  robotics research  is  the generation  of
dynamically balanced walk patterns. Since  the introduction of  the ZMP
formalism  \cite{vukobratovic1969contribution},  several  methods  have  been  proposed  to
generate  walking  motions efficiently.   One  way  to  deal with  the
complexity  of a  humanoid  robot  kinematic tree  is  to use  the
so-called "cart-table" simplified model \cite{kajita2003biped}. Based on this
model,  planning a trajectory  for the  ZMP is  reduced to  planning a
trajectory  for  the Center  of  Mass (CoM)  of  the  robot.  Given  a
trajectory  of  the CoM  and  footprint  positions, inverse  kinematics
solvers can animate  the whole set of DoFs of the
robot to generate a dynamically balanced walk trajectory.

@@

\section{From Statically Balanced Paths to Dynamic Walk Trajectories}
\label{sec:wb-step}

@@

\paragraph{Use of ZMP preview controller:} The control strategy presented in the  
previous proof may
generate very  long trajectories, because  of the transient  states at
the beginning and end of the locomotion. In the actual implementation,
we have chosen  to generate  CoM motions with  a ZMP preview  controller, as
presented in \cite{kajita2003biped}.  We  have observed experimentally that the
amplitude of  CoM trajectories decreases  when the frequency  of steps
increases. Our current ZMP preview controller relies on the cart-table model approximation.
To make this approximation valid, we fix the height of the robot CoM during walk,
as well as the vertical orientation of the robot waist. These geometric constraints 
are also applied when planning statically balanced paths, to ensure that the paths
can be approximated by dynamic walk trajectories. Note that this is
due to our current  ZMP preview controller implementation, and does not affect
the generality of the small-space controllability result presented above.

@@

\subsection{Application: Dynamic Approximation of a Statically Balanced Sliding Path}
\label{sec:ssc-application}

The  algorithm   that  animates  a  statically  balanced   path  into  a
dynamically balanced  walk trajectory has been inspired  by the previous
small-space controllability  proof.
Given a statically  balanced path $p$ respecting the cart-table model approximation constraints,
we  start  by placing  footprints  corresponding  to  the nominal  walk
pattern  of the  robot. Given the footprints, we  compute a ZMP trajectory, 
and a preview controller outputs  a corresponding CoM trajectory. 
The stack of tasks applied to
the robot to generate a dynamic walk motion is -- in decreasing priority order:

@@

Tasks (1)  and (2) generate a  dynamically balanced motion  by using the
simplified cart-table model  and the ZMP criterion. Tasks  (3) and (4)
ensure that the  resulting motion is well described  by the cart-table
model. Task (5)  is used to approximate $p$ as  well as possible given
the walk parameters.
